---
title: "Seguro Popular Analysis"
author: "Health Policy Analyst"
format: html
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(gt)
library(marginaleffects)
```

```{r}
#| cache: true
x <- sps |> 
  select(treatment, t2_health_exp_3m, age, sex, education) |> 
  drop_na() |> 
  mutate(treatment = factor(if_else(treatment == 1, "Enrolled", "Not enrolled"),
                            levels = c("Not enrolled", "Enrolled")))
```

```{r}
#| cache: true
fit_sp <- linear_reg(engine = "lm") |> 
  fit(t2_health_exp_3m ~ treatment, data = x)
```

```{r}
pred <- plot_predictions(fit_sp, condition = "treatment", draw = FALSE)

ggplot(pred, aes(x = treatment, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange(color = "#2b5c8f", size = 0.8) +
  theme_minimal() +
  labs(
    title = "Out-of-Pocket Health Spending by Seguro Popular Enrollment",
    subtitle = "Enrolling in Seguro Popular is associated with lower average out-of-pocket medical spending",
    caption = "Data Source: Mexico Seguro Popular Evaluation Experiment (King et al., 2009)",
    x = "Seguro Popular Enrollment Status",
    y = "Estimated 3-Month Out-of-Pocket Spending (Pesos)"
  )
```

Evaluating public health insurance programs is essential for understanding whether safety-net policies effectively protect uninsured families from financial hardship. Using household survey data from Mexico's Seguro Popular experiment, we examine whether enrolling in public health insurance reduces out-of-pocket medical expenditure over a three-month period. We model three-month out-of-pocket health spending as a linear function of household enrollment status in Seguro Popular. However, a key limitation of our model is that heavy-tailed health expenditure data with extreme outliers can disproportionately skew mean effect estimates. We estimate that enrollment in Seguro Popular reduces three-month out-of-pocket health spending by an average of 63 pesos, with a 95% confidence interval ranging from a 105-peso to a 22-peso reduction.
