---
title: "Seguro Popular Analysis"
author: "Health Policy Analyst"
format: html
execute:
  echo: false
---

Evaluating public health insurance programs is essential for understanding whether safety-net policies effectively protect uninsured families from financial hardship. Using household survey data from Mexico's Seguro Popular experiment, we examine whether enrolling in public health insurance reduces out-of-pocket medical expenditure over a three-month period.

```{r}
#| message: false
library(tidyverse)
library(primer.data)
```

```{r}
#| cache: true
x <- sps |> 
  select(treatment, t2_health_exp_3m, age, sex, education) |> 
  drop_na() |> 
  mutate(treatment = factor(if_else(treatment == 1, "Enrolled", "Not enrolled"),
                            levels = c("Not enrolled", "Enrolled")))
```

```{r}
ggplot(x, aes(x = treatment, y = t2_health_exp_3m)) +
  geom_jitter(alpha = 0.2, width = 0.2) +
  stat_summary(fun = mean, geom = "point", color = "red", size = 3) +
  coord_cartesian(ylim = c(0, 5000)) +
  labs(
    x = "Treatment Status",
    y = "Out-of-Pocket Spending (Pesos, 3 Months)",
    title = "Out-of-Pocket Health Spending by Treatment Group"
  )
```

```{r}
x |> count(treatment)
```


